### 2017-04-24 Inelastic/inelastic_main.tex

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.1 distribution.
%   Version 4.1r of REVTeX, August 2010
%
%   Copyright (c) 2009, 2010 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.1
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
 reprint,
superscriptaddress,
linenumbers,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%showpacs,preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps,
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}
%]{article}

\usepackage[normalem]{ulem}
\usepackage{color}
\usepackage{graphicx}% Include figure files
\usepackage[raggedright]{subfigure}
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{multicol}
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

%\usepackage[showframe,%Uncomment any one of the following lines to test 
%%scale=0.7, marginratio={1:1, 2:3}, ignoreall,% default settings
%%text={7in,10in},centering,
%%margin=1.5in,
%%total={6.5in,8.75in}, top=1.2in, left=0.9in, includefoot,
%%height=10in,a5paper,hmargin={3cm,0.8in},
%]{geometry}

%\usepackage[sorting=none]{biblatex}

\newcommand{\bern}{\affiliation{Albert Einstein Center for Fundamental Physics, University of Bern, 3012 Bern, Switzerland}}

\newcommand{\bologna}{\affiliation{Department of Physics and Astrophysics, University of Bologna and INFN-Bologna, 40126 Bologna, Italy}}

\newcommand{\chicago}{\affiliation{Department of Physics \& Kavli Institute of Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}}

\newcommand{\coimbra}{\affiliation{LIBPhys, Department of Physics, University of Coimbra, 3004-516 Coimbra, Portugal}}

\newcommand{\columbia}{\affiliation{Physics Department, Columbia University, New York, NY 10027, USA}}

\newcommand{\lngs}{\affiliation{INFN-Laboratori Nazionali del Gran Sasso and Gran Sasso Science Institute, 67100 L'Aquila, Italy}}

\newcommand{\mainz}{\affiliation{Institut f\"ur Physik \& Exzellenzcluster PRISMA, Johannes Gutenberg-Universit\"at Mainz, 55099 Mainz, Germany}}

\newcommand{\heidelberg}{\affiliation{Max-Planck-Institut f\"ur Kernphysik, 69117 Heidelberg, Germany}}

\newcommand{\munster}{\affiliation{Institut f\"ur Kernphysik, Westf\"alische Wilhelms-Universit\"at M\"unster, 48149 M\"unster, Germany}}

\newcommand{\nikhef}{\affiliation{Nikhef and the University of Amsterdam, Science Park, 1098XG Amsterdam, Netherlands}}

\newcommand{\nyuad}{\affiliation{New York University Abu Dhabi, Abu Dhabi, United Arab Emirates}}

\newcommand{\purdue}{\affiliation{Department of Physics and Astronomy, Purdue University, West Lafayette, IN 47907, USA}}

\newcommand{\rpi}{\affiliation{Department of Physics, Applied Physics and Astronomy, Rensselaer Polytechnic Institute, Troy, NY 12180, USA}}

\newcommand{\rice}{\affiliation{Department of Physics and Astronomy, Rice University, Houston, TX 77005, USA}}

\newcommand{\stockholm}{\affiliation{Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, Stockholm SE-10691, Sweden}}

\newcommand{\subatech}{\affiliation{SUBATECH, IMT Atlantique, CNRS/IN2P3, Universit\'e de Nantes, Nantes 44307, France}}

\newcommand{\torino}{\affiliation{INFN-Torino and Osservatorio Astrofisico di Torino, 10125 Torino, Italy}}

\newcommand{\ucla}{\affiliation{Physics \& Astronomy Department, University of California, Los Angeles, CA 90095, USA}}

\newcommand{\ucsd}{\affiliation{Department of Physics, University of California, San Diego, CA 92093, USA}}

\newcommand{\wis}{\affiliation{Department of Particle Physics and Astrophysics, Weizmann Institute of Science, Rehovot 7610001, Israel}}

\newcommand{\zurich}{\affiliation{Physik-Institut, University of Zurich, 8057  Zurich, Switzerland}}

\newcommand{\paris}{\affiliation{LPNHE, Université Pierre et Marie Curie, Université Paris Diderot, CNRS/IN2P3, Paris 75252, France}}

\newcommand{\freiburg}{\affiliation{Physikalisches Institut, Universität Freiburg, 79104 Freiburg, Germany}}

\begin{document}

\title{Search for WIMP Inelastic Scattering Off Xenon Nuclei With XENON100}% Force line breaks with \\
\author{E.~Aprile}\columbia

\author{J.~Aalbers}\nikhef

\author{F.~Agostini}\lngs\bologna

\author{M.~Alfonsi}\mainz

\author{F.~D.~Amaro}\coimbra

\author{M.~Anthony}\columbia

\author{F.~Arneodo}\nyuad

\author{P.~Barrow}\zurich

\author{L.~Baudis}\zurich

\author{B.~Bauermeister}\stockholm

\author{M.~L.~Benabderrahmane}\nyuad

\author{T.~Berger}\rpi

\author{P.~A.~Breur}\nikhef

\author{A.~Brown}\nikhef

\author{E.~Brown}\rpi

\author{S.~Bruenner}\heidelberg

\author{G.~Bruno}\lngs

\author{R.~Budnik}\wis

\author{L.~B\"utikofer}\freiburg\altaffiliation[]{Also at Albert Einstein Center for Fundamental Physics, University of Bern, Bern, Switzerland}

\author{J.~Calv\'en}\stockholm

\author{J.~M.~R.~Cardoso}\coimbra

\author{M.~Cervantes}\purdue

\author{D.~Cichon}\heidelberg

\author{D.~Coderre}\freiburg\altaffiliation[]{Also at Albert Einstein Center for Fundamental Physics, University of Bern, Bern, Switzerland}

\author{A.~P.~Colijn}\nikhef

\author{J.~Conrad}\altaffiliation{Wallenberg Academy Fellow}\stockholm

\author{J.~P.~Cussonneau}\subatech

\author{M.~P.~Decowski}\nikhef

\author{P.~de~Perio}\columbia

\author{P.~Di~Gangi}\bologna

\author{A.~Di~Giovanni}\nyuad

\author{S.~Diglio}\subatech

\author{G.~Eurin}\heidelberg

\author{J.~Fei}\ucsd

\author{A.~D.~Ferella}\stockholm

\author{A.~Fieguth}\munster

\author{W.~Fulgione}\lngs\torino

\author{A.~Gallo Rosso}\lngs

\author{M.~Galloway}\zurich

\author{F.~Gao}\columbia

\author{M.~Garbini}\bologna

\author{C.~Geis}\mainz

\author{L.~W.~Goetzke}\columbia

\author{Z.~Greene}\columbia

\author{C.~Grignon}\mainz

\author{C.~Hasterok}\heidelberg

\author{E.~Hogenbirk}\nikhef

\author{R.~Itay}\wis

\author{B.~Kaminsky}\freiburg\altaffiliation[]{Also at Albert Einstein Center for Fundamental Physics, University of Bern, Bern, Switzerland}

\author{S.~Kazama}\zurich

\author{G.~Kessler}\zurich

\author{A.~Kish}\zurich

\author{H.~Landsman}\wis

\author{R.~F.~Lang}\purdue

\author{D.~Lellouch}\wis

\author{L.~Levinson}\wis

\author{Q.~Lin}\columbia

\author{S.~Lindemann}\heidelberg\freiburg

\author{M.~Lindner}\heidelberg

\author{F.~Lombardi}\ucsd

\author{J.~A.~M.~Lopes}\altaffiliation[Also with ]{Coimbra Engineering Institute, Coimbra, Portugal}\coimbra

\author{A.~Manfredini}\wis

\author{I.~Maris}\nyuad

\author{T.~Marrod\'an~Undagoitia}\heidelberg

\author{J.~Masbou}\subatech

\author{F.~V.~Massoli}\bologna

\author{D.~Masson}\purdue

\author{D.~Mayani}\zurich

\author{M.~Messina}\columbia

\author{K.~Micheneau}\subatech

\author{A.~Molinario}\lngs

\author{K.~Mora}\stockholm

\author{M.~Murra}\munster

\author{J.~Naganoma}\rice

\author{K.~Ni}\ucsd

\author{U.~Oberlack}\mainz

\author{P.~Pakarha}\zurich

\author{B.~Pelssers}\stockholm

\author{R.~Persiani}\subatech

\author{F.~Piastra}\zurich

\author{J.~Pienaar}\purdue

\author{V.~Pizzella}\heidelberg

\author{M.-C.~Piro}\rpi

\author{G.~Plante}\columbia

\author{N.~Priel}\wis

\author{L.~Rauch}\heidelberg

\author{S.~Reichard}\purdue

\author{C.~Reuter}\purdue

\author{A.~Rizzo}\columbia

\author{S.~Rosendahl}\munster

\author{N.~Rupp}\heidelberg

\author{J.~M.~F.~dos~Santos}\coimbra

\author{G.~Sartorelli}\bologna

\author{M.~Scheibelhut}\mainz

\author{S.~Schindler}\mainz

\author{J.~Schreiner}\heidelberg

\author{M.~Schumann}\freiburg

\author{L.~Scotto~Lavina}\paris

\author{M.~Selvi}\bologna

\author{P.~Shagin}\rice

\author{M.~Silva}\coimbra

\author{H.~Simgen}\heidelberg

\author{M.~v.~Sivers}\freiburg\altaffiliation[]{Also at Albert Einstein Center for Fundamental Physics, University of Bern, Bern, Switzerland}

\author{A.~Stein}\ucla

\author{D.~Thers}\subatech

\author{A.~Tiseni}\nikhef

\author{G.~Trinchero}\torino

\author{C.~Tunnell}\nikhef\chicago

\author{M.~Vargas}\munster

\author{H.~Wang}\ucla

\author{Z.~Wang}\lngs

\author{Y.~Wei}\zurich

\author{C.~Weinheimer}\munster

\author{J.~Wulf}\zurich

\author{J.~Ye}\ucsd

\author{Y.~Zhang.}\columbia

\collaboration{XENON Collaboration}\email{xenon@lngs.infn.it}\noaffiliation

\date{\today}

\begin{abstract}
We present the first constraints on the spin-dependent, inelastic scattering cross section of Weakly Interacting Massive Particles (WIMPs) on nucleons from XENON100 data with an exposure of 7.64$\times$10$^3$\,kg\,day. XENON100 is a dual-phase xenon time projection chamber with 62\,kg of active mass, operated at the Laboratori Nazionali del Gran Sasso (LNGS) in Italy and designed to search for nuclear recoils from WIMP-nucleus interactions. Here we explore inelastic scattering, where a transition to a low-lying excited nuclear state of $^{129}$Xe is induced. The experimental signature is a nuclear recoil observed together with the prompt de-excitation photon. We see no evidence for such inelastic WIMP-$^{129}$Xe interactions. A profile likelihood analysis allows us to set a 90\% C.L. upper limit on the inelastic, spin-dependent WIMP-nucleon cross section of $3.3 \times 10^{-38}$\,cm$^{2}$  at 100\,GeV/c$^2$.  This is the most constraining result to date, and sets the pathway for an analysis of this interaction channel in upcoming, larger dual-phase xenon detectors.

\end{abstract}

%\keywords{Suggested keywords}%Use showkeys class option if keyword
                              %display desired
\maketitle

%\tableofcontents

\input{intro}

\input{analysis}

\input{results}

%\input{appendix}

\bibliography{inelastic_main} 
\bibliographystyle{unsrt}
%\printbibliography

\end{document}