### 2017-06-11 Inelastic/response_to_referee.tex
addedd response to referee

\documentclass{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}

\usepackage{color}

\begin{document}

\title{ \textbf{ Search for WIMP inelastic scattering off xenon nuclei with XENON100} \\  Response on Referee report -- DE12171/Aprile  }
%\author{}

\maketitle

%\section{Introduction}
%Here is the text of your introduction.

\begin{itemize}
	\item {\em{General: are these new SD limits isospin independent, applying to both protons and neutrons, unlike the traditional elastic SD for dark matter? If so, should be indicated right off the bat in the abstract.}}
\end{itemize}

\begin{itemize}
	\item{ \em{Introduction, third paragraph: Can you add a sentence further
elucidating how it is possible for dark matter, of the masses and
cross-sections you probe in this paper, to fail to be detected through
elastic scattering, given current existing world-best limits from
XENON1T and LUX and PandaX, yet still be possible through inelastic?
It was still unclear to me. Basically, I want to understand if there
was any reasonable scenario under which you could’ve seen a signal, or
whether that was already impossible and you were “doomed” to only set
a limit. This changes the theme of the paper significantly in my mind
from discovery potential to a confirmation/cross-check of an existing
limit in a new channel. - “10 keV of deposited energy”: Does this
refer to nuclear or electron recoil or both? Unclear since we are now
dealing with inelastic scattering that produces a gamma too here.}}
\end{itemize}

\textcolor{red}{ We do explicitly say that this search is not competitive, but complementary, maybe express that better?}
Classical searches for spin dependent elastic scattering have, and will keep to have a 
higher sensitivity with respect to this search for spin dependent inelastic scattering.
However, in case dark matter would be found, this type of search ensures you that the type 
{\color{red} of interaction is spin-dependent, while for classical elastic scattering is much more difficult to tell.}

\begin{itemize}
	\item{ \em{
Section III: Is there a reason the complete 477 live-days cannot be
used for this paper? Wouldn’t this make your constraints even
stronger? Please explain reason}}
\end{itemize}

Yes, the limit would be stronger. However the scope of this paper is 
not to discover dark matter, for which one would use instead SD elastic
scattering channel, but rather explore a different analysis channel.
Doing a combination on three different science runs (each of them having 
different background conditions and acceptances) is very demandig.

\begin{itemize}
	\item \em{
III.B: How were the ROI limits chosen? Text doesn’t say: should
explain, are these +/- some number of sigma relative to the expected
means of the NR and ER in cS1 and cS2 space?}
\end{itemize}

{\color{red} The signal distribution changes according to wimp mass due to the NR component,
we choose the ROI so that it will embrace all the wimp hipothesys at least two 
sigma. Now added to the text.}

\begin{itemize}
	\item \em{
While I understand the sub-regions were chosen to have roughly equal
number of events, it is not very clear to me at all why there are so
many different ones, and why there seems to be no rhyme or reason to
their sizes/shapes. It is also unclear, especially since Figure 1a is
an example of only one mass (100 GeV WIMP) how these regions were
picked to be optimal for any WIMP mass studied, doesn’t seem to be
possible to me. (Lastly, why didn’t you just take equal slices in S1
and S2 and make squares for example?)}
\end{itemize}

It was shown ({\color{red} see figure} ) a non appreciable
difference in sensitivity dividing the region in more than 10 equal sized bins. 
So provided that bins have similar sizes their shape is irrelevant to the sensitivity.
There is a limiting factor tough, we are using asymptotic formulae for distribution of the
test statistic, these formulaes are valid in the assumption of Wilks theorem, so that the number
of events in each bin should "large". In order to not have bins with drammaticly low background, which
happens in case of equally sized bins, we opted for collapsing bins achieving similar background content. 
This reasults in a somewhat arbitrary choise for some region of the ROI, but it does not impact our sensitivity.

\begin{itemize}
	\item \em{
Multiple-scatter cut: While I understand the cut on a second largest
peak, what happens when the peaks from multiple scatters are merged
because they are too close together? Also, the language is slightly
unclear: does “regardless of S2 signal size” mean the primary S2 peak,
or all peaks added together in a given event, counting them all as
“the signal”?}
\end{itemize}

{\color{red} When multiple scatters are closer than few mm the they will be mearged. don't know. }

\begin{itemize}
	\item \em{
Figure 2 caption slightly unclear: I assume the figures include both
the ER and the NR associated with the inelastic scatter, and not just
the former as the text implies. Please correct, to better match up
with explanation below in Section C.}
\end{itemize}

{\color{red} Yes, indeed. Text has been edited. }

\begin{itemize}
	\item \em{
Equation 2: It is not clear where S1 photon detection efficiency gets
folded in here: currently this equation appears to yield photons not
PE (Ly?) Am I reading it incorrectly?}
\end{itemize}

{\color{red} need to check. I forgot.}

\begin{itemize}
	\item \em{
~0.80 +/- 0.05: Given the importance of knowing the final signal
acceptance fraction I don’t understand why this has a “~,” especially
since error bars are provided. Why is this still approximate? It
should not be, unless you meant it changes with WIMP mass, and in that
case a range should be provided, from the min to max masses you
considered.}
\end{itemize}

What we meant is that is approximately constant with the wimp mass.
Values being compatible within uncertainties with 0.80 +- 0.05.

\begin{itemize}
	\item \em{
Section D: seems odd, doesn’t seem to add much beyond C and Figure 2.
If there is no new information here, please remove it, or at reduce
the length of the paragraph to include only those, and merge with C.
Weird to have a section be one short paragraph by itself.}
\end{itemize}

\begin{itemize}
	\item \em{
E: It is not clear to me how 60Co and 232Th calibrations can tell you
the BG in the signal region, given they have different energy
spectra/distributions than BG? Isn’t it better to validate the
simulation for 60Co and 232Th then simulate using NEST for example
events of the same energy distribution as your actual ER BG? In fact,
the BG simulation here seems to contradict the signal simulation in
the earlier section. Why the dichotomy in methods, especially since
the 39.6 keV line was actually present in calibration data, and so
this all seems backwards to me. This could explain your unresolved
systematic in F. you write is outside statistical fluctuation
expectation. (You should have a BG prior from counting the parts of
XENON100 and simulating in Geant4, yes?)}
\end{itemize}

The suggested approach relies on ..., there is no such a proven MC-data
matching for XENON100, meaning we don't have a Geant4 of all radioactive material 
background that produces a spectra that we can fully trust... We opted instead
for a simpler approach, bla bla bla....

\begin{itemize}
	\item \em{
Figure 3: statistical and systematic uncertainties should be denoted
separately as is more standard practice, either splitting them off or
using color coded error bars, as you do later}
\end{itemize}

\begin{itemize}
	\item \em{Pseudo-samples: I am confused as to why these are necessary, doesn’t
the ROOT PLR function allow profiling out over multiple such
systematic uncertainties simultaneously and continuously, and not just
in discrete +/-1-sigma steps? You also should explain whether
parameters are all railed to their plus/minus 1-sigma’s all together
or separately? This could be over-conservative, or swing the other
way, depending on the correlations on these parameters, which I know
for a fact myself can be very tricky: for example, raising a light
yield can be canceled out by lowering light collection efficiency,
each within its respective uncertainties, and not necessarily in
integer multiples of their standard deviations of course. This sounds
to me like what would be done in an “enhanced” cut-and-count-style
analysis, but not in a PLR. Perhaps it is because you did binned and I
am used to using PLR continuous (unbinned)? Can you please clarify,
and justify your choice of binned versus unbinned? This goes back to
my original question about the 9 strange sub-regions: How, a priori,
were they selected, optimized, and validated? How does one choose
one’s bins? Was it blind? Speaking of which, it should be made clear
what was blind or not in the text. I imagine not much since you are
primarily an NR-search experiment and this has ER Bottom of p. 6: How
can all parameters related to systematics be justified as normal, as
opposed to Poisson or skewed-normal in at least some cases? What I
mean is, for example a low-energy NR with a very low expected number
of photons detected is not going to have a symmetric distribution,
since there is no such thing as a negative number of photons. Does not
apply to higher energies, but I thought you have low O(10 keV) from
the NR here too Figure 4: minor comment, but sentences just a bit out
of order. Last sentence should move earlier because it refers to
discussion about the top half of the figure.}
\end{itemize}

{color{red} not really sure what's the issue here, several issues in one question....}
The analysis was not technically blind, since this data region was explored in several 
other studies. The procedure we used was as if it was blind, since we didn't tune cuts on data.

\begin{itemize}
	\item \em{Section IV: Please briefly define in text the difference between CL(s)
and the more common CL, and add a sentence explaining the benefit of
CL(s), or at least a brief reason why it was used instead of CL. Not
all readers may be familiar with Read’s work.}
\end{itemize}

\begin{itemize}
	\item \em{
Figure 5: Was the result constrained, due to large BG
under-fluctuation, to be at -1-sigma-level, as it appears to be in
this figure? If so, this should be stated in the figure caption and/or
text and explained. Also, based on flat regions I see in the 1- and
2-sigma bands in the plot it appears as though the limit was smoothed,
but the bands were not. Both should be left as raw data points, or
both smoothed, please choose, but be consistent (personally I prefer
smooth, but a definite case can be made for discrete connected by
straight lines in log-log in between the individual mass points you
actually run in your PLR; past XENON and LUX and other results vary in
choice of what to do in a case like this). This discrepancy is most
noticeable between 100 and 200 GeV, where the blue bands are extremely
straight, while the limit is curved. Lastly, are there really no other
results to put on this plot? Not even an earlier XENON one? This is
very surprising, but it does underscore the significance of your work
here, as being in nearly virgin territory, and you can say that.}
\end{itemize}

\end{document}