### 2018-08-17 SR1_detailed_note/Sections/DetectorSignalResponse.tex
SR1

XENON1T WIMP signal and background model used in the statistical inference are built based on the simulation which takes into consideration the detailed characterizations of the detector systematics, as well as the most comprehensible modeling of the recombination process of the liquid xenon.
The coherent modeling of the electronic recoils (ERs) and nuclear recoils (NR) is of high accuracy in terms of the uncertainty of the nuisance parameters and their correlations.
In this chapter, the details of the detector signal response model will be given.

\subsection{Basic Signal Response in Liquid Xenon}
% @Qing

\subsection{Recombination in Low Energy}
% @Qing

\subsection{Influence by Detector Reconstruction}
% @Qing