### 2018-08-17 SR1_detailed_note/bibliography.tex

\begin{thebibliography}{24}
\addcontentsline{toc}{chapter}{Bibliography}

\bibitem{wimp_hooper} G.~Bertone, D.~Hooper and J.~Silk, Phys. Rep. \textbf{405}, 279 (2005).

\bibitem{wimp_review} L.~Roszkowski {\it et al.} Rep. Prog. Phys. \textbf{81}, 066201 (2018). %[arXiv:1707.06277]

\bibitem{teresa_review} T.~Marrod\'an Undagoitia and L.~Rauch, J.\ Phys.\ G {\bf 43}, no. 1, 013001 (2016)

\bibitem{indirect} L.~E.~Strigari, Phys. Rep. \textbf{531}, 1 (2012).
%6
\bibitem{xe_sr0} E.~Aprile {\it et al.} (XENON Collaboration), Phys. Rev. Lett. \textbf{119}, 181301 (2017). %[arXiv:1705.06655].
%7
\bibitem{lux_dm} D. S. Akerib {\it et al.} (LUX Collaboration), Phys. Rev. Lett. \textbf{118}, 021303 (2016). %[arXiv:1608.07648].
%8
\bibitem{panda_dm} X. Cui {\it et al.} (PandaX-II Collaboration), Phys. Rev. Lett. \textbf{119}, 181302 (2017). %[arXiv:1708.06917].
%9
\bibitem{xe_instrument} E.~Aprile {\it et al.} (XENON Collaboration), Eur. Phys. J. C \textbf{77}: 881 (2017). %[arXiv:1708.07051].
\bibitem{xe_pmts} E.~Aprile {\it et al.} (XENON Collaboration), Eur. Phys. J. C \textbf{75}: 546 (2015). %[arXiv:1503.07698]. 
\bibitem{pmt2} P. Barrow {\it et al.} JINST \textbf{12}, no. 01, P01024 (2017). % [arXiv:1609.01654].

%10
\bibitem{xe_muonveto} E.~Aprile {\it et al.} (XENON Collaboration), JINST \textbf{9}, P11006 (2014). %[arxiv:1406.2374].

\bibitem{sorensen} P.~Sorensen, K.~Kamdin. JINST \textbf{13}, P02032 (2018). %[arxiv:1711.07025].

\bibitem{kemfield} D.~Furse {\it et al.}. New J. Phys. \textbf{19} 053012 (2017)

\bibitem{luxkr} D. S. Akerib {\it et al.}, Phys. Rev. D \textbf{96}, 112009 (2017). % [arXiv:1708.02566].

% rn220
\bibitem{xe_rn220} E.~Aprile {\it et al.} (XENON Collaboration), Phys. Rev. D \textbf{95}, 072008 (2017).

%11
\bibitem{xe1t_ng} R.~F.~Lang {\it et al.}, Nucl. Inst. and Meth. A \textbf{879}, 31 (2018).

\bibitem{pearc_computing} B. Riedel {\it et al.} PEARC '18, ISBN 978-1-4503-6446-1 (2018). doi:10.1145/3219104.3219155.

\bibitem{pax} XENON Collaboration. (2018). The pax data processor v6.8.0. Zenodo. http://doi.org/10.5281/zenodo.1195785
%14
\bibitem{pmt_calibration} R.~Saldanha {\it et al.}, Nucl. Inst. and Meth. A \textbf{863}, 35 (2017). %[arXiv:1602.03150].
%15
\bibitem{double_pe} C. H. Faham {\it et al.}, JINST \textbf{10}, P09010 (2015).

\bibitem{lux_field} D. S. Akerib {\it et al.} (LUX Collaboration), JINST \textbf{12}, P11022 (2017).

\bibitem{lewin} J. Lewin and P. Smith, Astropart. Phys. \textbf{6} 87 (1996).

\bibitem{xe_kr} E.~Aprile {\it et al.} (XENON Collaboration), Eur. Phys. J. C \textbf{77}: 275 (2017). %[arXiv:1612.04284].
%16
\bibitem{xe_sebastian} S. Lindemann, H. Simgen, Eur. Phys. J. C \textbf{74}, 2746 (2014). %[arXiv:1308.4806].

\bibitem{xe_rn} E.~Aprile {\it et al.} (XENON Collaboration),	Eur. Phys. J. C \textbf{78} :132 (2018). %[arXiv:1708.03617]

%17
\bibitem{solarflux} A.~M.~Serenelli {\it et al.}, Astro. Phys. Journal \textbf{743}, 29 (2011). %[arXiv:1104.1639].
%18
\bibitem{coherent} D.~Akimov {\it et al.} (COHERENT Collaboration), Science \textbf{357}, 1123-1126 (2017). %[arXiv:1708.01294].

\bibitem{geant4} S.~Agostinelli {\it et al.}, Nucl. Inst. and Meth. A \textbf{506}, 250 (2003).

\bibitem{xenon1t_mc_paper} E.~Aprile {\it et al.} (XENON Collaboration), J.\ Cosmol.\ Astropart.\ Phys. \textbf{1604}, no. 04, 027 (2016)

\bibitem{xe_screen} E.~Aprile {\it et al.} (XENON Collaboration), Eur. Phys. J. C \textbf{77}: 890 (2017).

\bibitem{sources4a} W.B.~Wilson {\it et al.}, LANL technical note LA-13639-MS (1999).

\bibitem{mcnpx} R.~Lemrani {\it et al.}, Nucl. Inst. and Meth. A \textbf{560}, 454 (2006). %[arXiv:0601030].

\bibitem{xe100_tritium} E.~Aprile {\it et al.} (XENON Collaboration), Phys. Rev. D \textbf{97}, 092007 (2018).

\bibitem{ti_model} J. Thomas, and D. A. Imel, Phys. Rev. A \textbf{36}, 614 (1987).

\bibitem{NEST_v1} B. Lenardo {\it et al.} IEEE Trans. Nucl. Sci. \textbf{62}, 3387 (2015).

\bibitem{lux_tritium} D. S. Akerib et al. (LUX Collaboration), Phys. Rev. D \textbf{93}, 072009 (2016).

\bibitem{pixey_ar37} E. M. Boulton {\it et al.}, JINST \textbf{12}, P08004 (2017).

\bibitem{lux_xe127} D. S. Akerib {\it et al.} (LUX Collaboration), Phys. Rev. D \textbf{96}, 112011 (2017).

\bibitem{feldmancousins} G.~J.~Feldman and R.~D.~Cousins, Phys. Rev. D {\bf 57}, 3873 (1998).

\bibitem{PDG} C.~Patrignani {\it et al.} (Particle Data Group), Chin. Phys. C \textbf{40}, 100001 (2016).

\bibitem{James:1980ci} F.~James, Comput.\ Phys.\ Commun.\  \textbf{ 20}, 29 (1980).

\bibitem{Bartlett:1953} M. S. Bartlett, Biometrika \textbf{40} 3/4 (1953).

\bibitem{safeguard} N. Priel, L. Rauch, H. Landsman, A. Manfredini, and R. Budnik, J.\ Cosmol.\ Astropart.\ Phys. \textbf{5} 13 (2017).

\bibitem{exo_nature} J.~B.~Albert {\it et al.} (EXO-200 Collaboration), Nature \textbf{510}, 225 (2014).

\end{thebibliography}