### 2018-08-17 SR1_detailed_note/main.tex

%\RequirePackage{lineno}\linenumbers\newdimen\linenumbersep\linenumbersep=2pt
\documentclass[prd,aps,twocolumn,superscriptaddress,showpacs,showkeys,amsmath,amssymb,floatfix]{revtex4}
\usepackage[colorlinks=true,citecolor=blue,filecolor=blue,linkcolor=blue,urlcolor=blue,pdftex]{hyperref}

\usepackage{xspace} 
\usepackage[usenames,dvipsnames]{color}
\usepackage{booktabs,graphicx,mathrsfs,verbatim,amsmath,units,soul}
\usepackage{xspace} 
\usepackage{upgreek}
\usepackage{gensymb}
\usepackage{amsmath}
\usepackage{tabularx}

\sloppy

% NUMBERS

%% **********************
%Intro
%% **********************
\newcommand{\ton}{t}
% total mass LXe
\newcommand{\nxenontotalmass}{$3.2$\;\ton\xspace}
% target mass LXe
\newcommand{\nxenontargetmass}{$2.00$\;\ton\xspace}
% distance cathode to gate
\newcommand{\ntpcdriftlength}{$97~\mathrm{cm}$\xspace}
% diameter of TPC
\newcommand{\ntpcdiameter}{$96~\mathrm{cm}$\xspace}
% extraction field in V/cm
% https://xe1t-wiki.lngs.infn.it/doku.php?id=xenon:xenon1t:julien:xenon1t_3d_field_simulation#s1-region_-_sr1
%  ^ Warning: seems incorrect...
\newcommand{\nextractionfield}{$\sim$4~kV/cm\xspace}
\newcommand{\nmultiplicationfield}{$\sim$8~kV/cm\xspace}
% X ER rejection at Y NR acceptance
\newcommand{\nerrejectionpct}{99.7\%\xspace}
\newcommand{\nnracceptance}{50\%\xspace}
\newcommand{\nxenonnttargetmass}{5.9\;\ton\xspace}

%% **********************
% Detector Parameters 
%% **********************
% Liquid level tolerance
\newcommand{\nliquidleveltolerance}{2\%~RMS\xspace}
% Liquid level above gate
\newcommand{\nliquidlevelabovegate}{$2.5~\mathrm{mm}$\xspace}
% LXe temperature
\newcommand{\nxenoninternaltemp}{$-96.0$\,$^\circ$C\xspace}
% LXe temperature tolerance
\newcommand{\nxenoninternaltemptolerance}{0.02\%~RMS\xspace}
% GXe pressure
\newcommand{\nxenonpressure}{1.94~bar\xspace}
% GXe pressure tolerance
\newcommand{\nxenonpressuretolerance}{0.01\%~RMS\xspace}
% Largest S2 x-y map correction (in middle)
\newcommand{\nlargeststwoxycorrection}{32\%\xspace}
% Drift Fields 
% https://xe1t-wiki.lngs.infn.it/doku.php?id=xenon:xenon1t:jingqiang:field:data_driven_3d_electric_field_map
\newcommand{\driftfieldsrzero}{$120$~V/cm\xspace} %\;\pm\;8
\newcommand{\driftfieldsrone}{$81$~V/cm\xspace} %\;\pm\;7
\newcommand{\driftfielderror}{2.2~V/cm RMS\xspace} %\;\pm\;7

%% **********************
% Livetimes
%% **********************
\newcommand{\SRzero}{SR0\xspace}
\newcommand{\SRone}{SR1\xspace}
% Livetime-corrected SR0
\newcommand{\nlivetimesrzero}{32.1~days\xspace}
% Livetime-corrected SR1
\newcommand{\nlivetimesrone}{246.7~days\xspace}
% Livetime-corrected combined
\newcommand{\nlivetimetotal}{278.8~days\xspace}
% Start date of sciencerun0
\newcommand{\nstartofsciencerunzero}{November~22,~2016\xspace}
% End date of sciencerun1
\newcommand{\nendofsciencerunone}{February~8,~2018\xspace}
% Power outage of sciencerun1
\newcommand{\npoweroutsciencerunone}{February~24,~2018\xspace}
% SR0 DAQ livetime
\newcommand{\ndaqlivetimesrzero}{†92.2\%\xspace}
\newcommand{\ndaqdeadtimesrzero}{7.8\%\xspace}
% SR1 DAQ livetime
\newcommand{\ndaqlivetimesrone}{98.8\%\xspace}
\newcommand{\ndaqdeadtimesrone}{1.2\%\xspace}
% Date of run-ending earthquake
\newcommand{\nearthquakedate}{January~18,~2017\xspace}
% Date SR1 start
\newcommand{\nsronestart}{February~2,~2017\xspace}
% Magnitude of run-ending earthquake
\newcommand{\nearthquakemagnitude}{5.7\xspace}
% Total gain zero PMTs
\newcommand{\nblindedpmts}{36\xspace}
% Gain zero PMTs due to leaks
\newcommand{\nleakypmts}{26\xspace}
% Gain zero PMTs not due to leaks (should add up to total)
\newcommand{\notherpmts}{6\xspace}
% Days lost from muon veto cut
\newcommand{\nlossmuon}{1.2\%\xspace}%{3.4~days\xspace}
% Days lost from PMT flash cut
\newcommand{\nlossflash}{0.2\%\xspace}%{0.6~days\xspace}
% Days lost from S2 tail cut
\newcommand{\nlosstails}{4.4\%\xspace}%{12.3~days\xspace}
% Days of good Rn220 data
\newcommand{\ndaysrn}{17.1 days\xspace}
% Days of good AmBe data
\newcommand{\ndaysambe}{30.0 days\xspace}
% Days of good NG data
\newcommand{\ndaysng}{1.9 days\xspace}

%% **********************
% Thresholds and acceptance
%% **********************
% S2 threshold above 200pe
\newcommand{\nstwoaccoverthresh}{unity\xspace}
% Mean SPE acceptance, 3% is the 1-sigma. RMS would be 9.2%
\newcommand{\nmeanspeacceptance}{$(93\pm3)$\%\xspace}

%% **********************
% Fiducial and backgrounds
%% **********************
% Fiducial mass
\newcommand{\ncoremassname}{0.65~\ton\xspace}
\newcommand{\nonetonnemassname}{0.9~\ton\xspace}
\newcommand{\nnominalfiducialmassname}{1.3~\ton\xspace}
\newcommand{\nnominalfiducialmass}{$(1.30~\pm~0.01)$\;\ton\xspace}
%\newcommand{\nfiducialmass}{$1.6$}
% Combined exposure
\newcommand{\nexposure}{1.0\;\ton$\times$yr\xspace}
% Total NR background in fiducial mass
\newcommand{\ntotalnrbackground}{$XX$}
% Total ER background in fiducial mass
\newcommand{\nwallbackground}{$XX$}
% Kr85 level
\newcommand{\nkrlevel}{$^\mathrm{nat}\mathrm{Kr}/\mathrm{Xe} = (0.66\pm0.11)~\mathrm{ppt}$}
\newcommand{\krrate}{$(7.7\pm 1.3)$~events/$(\mathrm{\ton}\times\mathrm{yr}\times\kevee)$}

\newcommand{\rnlevel}{$(12.6 \pm 0.8)~\mu\mathrm{Bq}/\mathrm{kg}$\xspace}

\newcommand{\rnrate}{$(7.7\pm 1.8)$~events/$(\mathrm{\ton}\times\mathrm{yr}\times\kevee)$}

%\newcommand{\nkrlevel}{$^{85}\mathrm{Kr}/^{nat}\mathrm{Kr} = (1.645 \substack{+0.413 \\ -0.276})\times 10^{-11}$}
% Rn222 level
\newcommand{\nrnlevel}{YY}
%\newcommand{\nlowestbg}{${(63~\pm~2)}$~\dru\xspace}
\newcommand{\nlowestbg}{$(82\substack{+5 \\ -3}\textrm{~(sys)}\pm3\textrm{~(stat)})$~\dru\xspace}

\newcommand{\rntwo}{${}^{222}$Rn\xspace}
\newcommand{\rnzero}{${}^{220}$Rn\xspace}
\newcommand{\krbkg}{${}^{85}$Kr\xspace}
\newcommand{\krm}{${}^{83\mathrm{m}}$Kr\xspace}
\newcommand{\ambe}{${}^{241}$AmBe\xspace}

\newcommand{\cSone}{$\mathrm{cS1}$\xspace}
\newcommand{\cStwo}{$\mathrm{cS2}$\xspace}
\newcommand{\cStwob}{$\mathrm{cS2_b}$\xspace}
\newcommand{\ucSone}{$\mathrm{S1}$\xspace}
\newcommand{\ucStwo}{$\mathrm{S2}$\xspace}
\newcommand{\Sone}{$\mathrm{S1}$\xspace}
\newcommand{\Stwo}{$\mathrm{S2}$\xspace}
\newcommand{\z}{Z\xspace}
\newcommand{\radius}{R\xspace}
\newcommand{\radiussquared}{R$^{2}$\xspace}

\newcommand{\nmoststringentlimit}{$4.1\times10^{-47}$~cm$^2$\xspace}
\newcommand{\nmoststingentmass}{30~\gevcsq}
\newcommand{\nbestfitsigmasi}{$2.9\times10^{-48}$~cm$^2$\xspace}
\newcommand{\nbestfitsigmasitwoh}{$4.7\times10^{-47}$~cm$^2$\xspace}

\newcommand\figref[1]{Fig.~\ref{#1}}
\newcommand\Figref[1]{Figure~\ref{#1}}
\newcommand\tabref[1]{Table~\ref{#1}}
\newcommand\figsref[2]{Figs.~\ref{#1}~and~\ref{#2}}
\newcommand\Figsref[2]{Figures~\ref{#1}~and~\ref{#2}}

\newcommand{\kevee}{\mathrm{keV_{ee}}}
\newcommand{\kevnr}{\mathrm{keV_{nr}}}

% Old abbreviations
\newcommand{\AC}{AC\xspace}
\newcommand{\FM}{fiducial mass\xspace}

% old microscopic units
%\newcommand{\dru}{($\mathrm{kg}\times\mathrm{day}\times\mathrm{keV})^{-1}$\xspace}
% new macroscopic units
\newcommand{\dru}{events/$(\mathrm{\ton}\times\mathrm{yr}\times\kevee)$}
\newcommand{\gevcsq}{GeV/c${}^2$\xspace}
\newcommand{\tevcsq}{TeV/c${}^2$\xspace}
\newcommand{\searchregion}{cS$1\!\in\![3, 70]\;\mathrm{PE}$, cS$2_b\!\in\![50, 8000]\;\mathrm{PE}$~search region}
% https://xe1t-wiki.lngs.infn.it/doku.php?id=greene:bbf_acceptances
\newcommand{\searchregionkev}{[1.4,~10.6]~$\kevee$\xspace}
\newcommand{\searchregionkevnr}{[4.9,~40.9]~$\kevnr$\xspace}

\newcommand{\todo}[1]{\textcolor{red}{\textbf{$\lceil$#1$\rfloor$}}}
\newcommand{\num}[1]{#1}

\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{\texttt{arXiv:#1}}}

\makeatletter
\newcommand{\romenum}[1]{(\romannumeral #1)}
\makeatother

\newcommand{\Delete}[2][magenta]{\textcolor{#1}{\sout{#2}}}
\newcommand{\Modify}[2][magenta]{\textcolor{#1}{\textrm{#2}}}

\newcommand{\bern}{\affiliation{Albert Einstein Center for Fundamental Physics, University of Bern, 3012 Bern, Switzerland}}

\newcommand{\bologna}{\affiliation{Department of Physics and Astronomy, University of Bologna and INFN-Bologna, 40126 Bologna, Italy}}

\newcommand{\chicago}{\affiliation{Department of Physics \& Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA}}

\newcommand{\coimbra}{\affiliation{LIBPhys, Department of Physics, University of Coimbra, 3004-516 Coimbra, Portugal}}

\newcommand{\columbia}{\affiliation{Physics Department, Columbia University, New York, NY 10027, USA}}

\newcommand{\lngs}{\affiliation{INFN-Laboratori Nazionali del Gran Sasso and Gran Sasso Science Institute, 67100 L'Aquila, Italy}}

\newcommand{\mainz}{\affiliation{Institut f\"ur Physik \& Exzellenzcluster PRISMA, Johannes Gutenberg-Universit\"at Mainz, 55099 Mainz, Germany}}

\newcommand{\heidelberg}{\affiliation{Max-Planck-Institut f\"ur Kernphysik, 69117 Heidelberg, Germany}}

\newcommand{\munster}{\affiliation{Institut f\"ur Kernphysik, Westf\"alische Wilhelms-Universit\"at M\"unster, 48149 M\"unster, Germany}}

\newcommand{\nikhef}{\affiliation{Nikhef and the University of Amsterdam, Science Park, 1098XG Amsterdam, Netherlands}}

\newcommand{\nyuad}{\affiliation{New York University Abu Dhabi, Abu Dhabi, United Arab Emirates}}

\newcommand{\purdue}{\affiliation{Department of Physics and Astronomy, Purdue University, West Lafayette, IN 47907, USA}}

\newcommand{\rpi}{\affiliation{Department of Physics, Applied Physics and Astronomy, Rensselaer Polytechnic Institute, Troy, NY 12180, USA}}

\newcommand{\rice}{\affiliation{Department of Physics and Astronomy, Rice University, Houston, TX 77005, USA}}

\newcommand{\stockholm}{\affiliation{Oskar Klein Centre, Department of Physics, Stockholm University, AlbaNova, Stockholm SE-10691, Sweden}}

\newcommand{\subatech}{\affiliation{SUBATECH, IMT Atlantique, CNRS/IN2P3, Universit\'e de Nantes, Nantes 44307, France}}

\newcommand{\torino}{\affiliation{INFN-Torino and Osservatorio Astrofisico di Torino, 10125 Torino, Italy}}

\newcommand{\ucla}{\affiliation{Physics \& Astronomy Department, University of California, Los Angeles, CA 90095, USA}}

\newcommand{\ucsd}{\affiliation{Department of Physics, University of California, San Diego, CA 92093, USA}}

\newcommand{\wis}{\affiliation{Department of Particle Physics and Astrophysics, Weizmann Institute of Science, Rehovot 7610001, Israel}}

\newcommand{\zurich}{\affiliation{Physik-Institut, University of Zurich, 8057  Zurich, Switzerland}}

\newcommand{\paris}{\affiliation{LPNHE, Universit\'{e} Pierre et Marie Curie, Universit\'{e} Paris Diderot, CNRS/IN2P3, Paris 75252, France}}

\newcommand{\freiburg}{\affiliation{Physikalisches Institut, Universit\"at Freiburg, 79104 Freiburg, Germany}}

\begin{document}

% CHANGE TO TITLE

\title{Background Modeling and Statistical Inference for WIMP Searches in XENON1T }

\input{authors.tex}

\date{\today}

\begin{abstract}

\end{abstract}

\pacs{
    95.35.+d, %Dark matter
    14.80.Ly, %Supersymmetric partners of known particles
    29.40.-n,  %Radiation detectors
    95.55.Vj
}

\keywords{Dark Matter, Direct Detection, Xenon, Data Analysis}

\maketitle

\section{Introduction}
% @all

\section{Detector Signal Response Model}
% @Qing

\input{Sections/DetectorSignalResponse.tex}

\section{Background and Dark Matter Signal Model}

\input{Sections/ERBackground.tex}

\input{Sections/NRBackground.tex}

\input{Sections/SurfaceBackground.tex}

\input{Sections/ACBackground.tex}

\input{Sections/WIMPSignalModel.tex}

\input{Sections/Inference.tex}
\subsection{Inference of XENON1T Data}
\input{Sections/Likelihood.tex}

\input{bibliography.tex}
\end{document}